---
title: "_The Economist_'s Presidential Election Forecasting Model"
subtitle: "Model report"
date: "Last update on `r format(Sys.time(), '%A %B %d, %Y at %I:%M %p %Z',tz='US/Eastern')`"
output: html_document
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(gridExtra)
library(boot)
library(pbapply)
library(urbnmapr)
library(ggrepel)
library(DT)

# final model version
# source('scripts/model/final_2016.R');source('scripts/model/final_2012.R');source('scripts/model/final_2008.R');rmarkdown::render('model_summary.Rmd',output_file = 'model_reports/v3_add_baseline_cov');beepr::beep(2);
```

## Model performance {.tabset .tabset-fade}

Overall:


Select an election for more:

### 2008 

```{r include=F}
## Master variables
RUN_DATE <- min(ymd('2008-11-03'),Sys.Date())

election_day <- ymd("2008-11-03")
start_date <- as.Date("2008-03-01") # Keeping all polls after March 1, 2016
  
evs_df <- read_csv('data/2012.csv')

# will read in the most recent back-test for 2008
out  <- read_rds(sprintf('models/backtest_2008/stan_model_%s.rds',RUN_DATE))

# get all the polling data
#setwd(here("data/"))
all_polls <- read_csv('data/all_polls_2008.csv')

# select relevant columns from HufFPost polls
all_polls <- all_polls %>%
  dplyr::select(state, pollster, number.of.observations, mode,population,
                start.date, 
                end.date,
                obama, mccain, undecided, other)%>%
  mutate(end.date = mdy(end.date),
         start.date = mdy(start.date))


# make sure we've got nothing from the futuree
all_polls <- all_polls %>%
  filter(ymd(end.date) <= RUN_DATE)

# basic mutations
df <- all_polls %>% 
  tbl_df %>%
  rename(n = number.of.observations) %>%
  mutate(begin = ymd(start.date),
         end   = ymd(end.date),
         t = end - (1 + as.numeric(end-begin)) %/% 2) %>%
  filter(t >= start_date & !is.na(t)
         & n > 1) 

# pollster mutations
df <- df %>%
  mutate(pollster = str_extract(pollster, pattern = "[A-z0-9 ]+") %>% sub("\\s+$", "", .),
         pollster = replace(pollster, pollster == "Fox News", "FOX"), # Fixing inconsistencies in pollster names
         pollster = replace(pollster, pollster == "WashPost", "Washington Post"),
         pollster = replace(pollster, pollster == "ABC News", "ABC"),
         undecided = ifelse(is.na(undecided), 0, undecided),
         other = ifelse(is.na(other), 0, other))

# vote shares etc
df <- df %>%
  mutate(two_party_sum = obama + mccain,
         polltype = as.integer(as.character(recode(population, 
                                                   "Likely Voters" = "0", 
                                                   "Registered Voters" = "1",
                                                   "Adults" = "2"))), 
         n_respondents = round(n),
         # obama
         n_obama = round(n * obama/100),
         pct_obama = obama/two_party_sum,
         # mccain
         n_mccain = round(n * mccain/100),
         pct_mccain = mccain/two_party_sum,
         # third-party
         n_other = round(n * other/100),
         p_other = other/100)


# state info
state_correlation <- read.csv("data/potus_results_76_16.csv") %>% 
    select(year, state, dem) %>%
    group_by(state) %>%
    mutate(dem = dem )  %>% 
    spread(state,dem) %>% 
    dplyr::select(-year)

#here("data")
state_data <- read.csv("data/potus_results_76_16.csv")

all_state_names <- unique(state_data$state)

# Numerical indices passed to Stan for states, days, weeks, pollsters
df <- df %>% 
  mutate(poll_day = t - min(t) + 1,
         # Factors are alphabetically sorted: 1 = --, 2 = AL, 3 = AK, 4 = AZ...
         index_s = as.numeric(factor(as.character(state),
                                     levels = c('--',all_state_names))),  # ensure levels are same as all 50 names in sate_correlation
         index_t = 1 + as.numeric(t) - min(as.numeric(t)),
         index_p = as.numeric(as.factor(as.character(pollster))))  

T <- as.integer(round(difftime(election_day, min(df$start.date))))

# selections
df <- df %>%
  arrange(state, t, polltype, two_party_sum) %>% 
  distinct(state, t, pollster, .keep_all = TRUE) %>%
  select(
    # poll information
    state, t, begin, end, pollster, polltype, method = mode, n_respondents, 
    # vote shares
    pct_obama, n_obama, 
    pct_mccain, n_mccain, 
    p_other, n_other, poll_day, index_s, index_p, index_t) %>%
  mutate(index_s = ifelse(index_s == 1, 52, index_s - 1)) # national index = 51


# state weights
#setwd(here("data/"))
state_elec_results <- read.csv("data/2008.csv", 
                       header = TRUE, stringsAsFactors = FALSE) %>% 
  mutate(score = obama_count / (obama_count + mccain_count),
         national_score = sum(obama_count)/sum(obama_count + mccain_count),
         delta = score - national_score,
         share_national_vote = (total_count)
         /sum(total_count)) %>%
  arrange(state) 

rownames(state_elec_results) <- state_elec_results$state

# get state incdices
all_states <- state_elec_results$state
state_name <- state_elec_results$state_name
names(state_name) <- state_elec_results$state

# set prior differences
prior_diff_score <- state_elec_results$delta
names(prior_diff_score) <- all_states

# set state weights
state_weights <- c(state_elec_results$share_national_vote / sum(state_elec_results$share_national_vote))
names(state_weights) <- c(state_elec_results$state)

# electoral votes, by state:
ev_state <- state_elec_results$ev
names(ev_state) <- state_elec_results$state

# hard code prior
national_mu_prior <- 0.566

# Mean of the mu_b_prior
mu_b_prior <- logit(national_mu_prior + prior_diff_score)

# etc
a <- rstan::extract(out, pars = "alpha")[[1]]
hist(a)
# sigmas
tibble(sigma_national = rstan::extract(out, pars = "sigma_a")[[1]],
       sigma_state = rstan::extract(out, pars = "sigma_b")[[1]]) %>%
  gather(parameter,value) %>%
  ggplot(.,aes(x=value)) +
  geom_histogram(binwidth=0.001) +
  facet_grid(rows=vars(parameter))
# measurement noise
measure_noise_national <- rstan::extract(out, pars = "measure_noise_national")[[1]] * mean(rstan::extract(out, pars = "sigma_measure_noise_national")[[1]])
hist(measure_noise_national)
# look at variation in mu_a
mu_a <- rstan::extract(out, pars = "mu_a")[[1]]
lapply(1:100,
       function(x){
         tibble(mu_a_draw = inv.logit(mu_a[x,]),
                trial = x) %>%
           mutate(date = min(df$end) + row_number()) 
       }) %>%
  do.call('bind_rows',.) %>%
  ggplot(.,aes(x=date,y=mu_a_draw,group=trial)) +
  geom_line(alpha=0.2)
# extract predictions
# states
predicted_score <- rstan::extract(out, pars = "predicted_score")[[1]]

pct_obama <- pblapply(1:dim(predicted_score)[3],
                    function(x){
                      # pred is mu_a + mu_b for the past, just mu_b for the future
                      temp <- predicted_score[,,x]
                      
                      # put in tibble
                      tibble(low = apply(temp,2,function(x){(quantile(x,0.05))}),
                             high = apply(temp,2,function(x){(quantile(x,0.95))}),
                             mean = apply(temp,2,function(x){(mean(x))}),
                             prob = apply(temp,2,function(x){(mean(x>0.5))}),
                             state = x) 
                      
                    }) %>% do.call('bind_rows',.)

pct_obama$state = colnames(state_correlation)[pct_obama$state]

pct_obama <- pct_obama %>%
  group_by(state) %>%
  mutate(t = row_number() + min(df$begin)) %>%
  ungroup()

# national
pct_obama_natl <- pblapply(1:dim(predicted_score)[2],
                         function(x){
                           # each row is a day for a particular draw
                           temp <- predicted_score[x,,] %>% as.data.frame()
                           names(temp) <- colnames(state_correlation)
                           
                           # for each row, get weigted natl vote
                           tibble(natl_vote = apply(temp,MARGIN = 1,function(y){weighted.mean(y,state_weights)})) %>%
                             mutate(t = row_number() + min(df$begin)) %>%
                             mutate(draw = x)
                         }) %>% do.call('bind_rows',.)

pct_obama_natl <- pct_obama_natl %>%
  group_by(t) %>%
  summarise(low = quantile(natl_vote,0.05),
            high = quantile(natl_vote,0.95),
            mean = mean(natl_vote),
            prob = mean(natl_vote > 0.5)) %>%
  mutate(state = '--')

# bind state and national vote
pct_obama <- pct_obama %>%
  bind_rows(pct_obama_natl) %>%
  arrange(desc(mean))

# look
ex_states <- c('IA','FL','OH','WI','MI','PA','AZ','NC','NH','TX','GA','MN')
pct_obama %>% filter(t == RUN_DATE,state %in% c(ex_states,'--')) %>% mutate(se = (high - mean)/1.68) %>% dplyr::select(-t)

map.2008.gg <- urbnmapr::states %>%
  left_join(pct_obama %>% filter(t == max(t)) %>%
              select(state_abbv=state,prob)) %>%
  ggplot(aes(x=long,y=lat,group=group,fill=prob)) +
  geom_polygon(col='gray40')  + 
  coord_map("albers",lat0=39, lat1=45) +
  scale_fill_gradient2(name='Democratic win probability',high='blue',low='red',mid='gray90',midpoint=0.5) +
  theme_void() + 
  theme(legend.position = 'top')


# electoral college by simulation
draws <- pblapply(1:dim(predicted_score)[3],
                  function(x){
                    # pred is mu_a + mu_b for the past, just mu_b for the future
                    pct_obama <- predicted_score[,,x]
                    
                    pct_obama <- pct_obama %>%
                      as.data.frame() %>%
                      mutate(draw = row_number()) %>%
                      gather(t,pct_obama,1:(ncol(.)-1)) %>%
                      mutate(t = as.numeric(gsub('V','',t)) + min(df$begin),
                             state = all_state_names[x]) 
                    
                    
                  }) %>% do.call('bind_rows',.)


sim_evs <- draws %>%
  left_join(state_elec_results %>% select(state,ev),by='state') %>%
  group_by(t,draw) %>%
  summarise(dem_ev = sum(ev * (pct_obama > 0.5))) %>%
  group_by(t) %>%
  summarise(mean_dem_ev = mean(dem_ev),
            median_dem_ev = median(dem_ev),
            high_dem_ev = quantile(dem_ev,0.975),
            low_dem_ev = quantile(dem_ev,0.025),
            prob = mean(dem_ev >= 270)) %>%
  left_join(pct_obama[pct_obama$state != '--',] %>%
              left_join(state_elec_results %>% select(state,ev),by='state') %>%
              group_by(t) %>%
              summarise(sum_dem_ev = sum(ev * (prob > 0.5))) )


# add identifier
identifier <- paste0(Sys.Date()," || " , out@model_name)

natl_polls.2008.gg <- pct_obama %>%
  filter(state == '--') %>%
  left_join(df %>% select(state,t,pct_obama)) %>% # plot over time
  # plot
  ggplot(.,aes(x=t)) +
  geom_ribbon(aes(ymin=low,ymax=high),col=NA,alpha=0.2) +
  geom_hline(yintercept = 0.5) +
  geom_hline(yintercept = national_mu_prior,linetype=2) +
  geom_point(aes(y=pct_obama),alpha=0.2) +
  geom_smooth(aes(y=pct_obama),method='loess',span=0.2,col='black',linetype=2,se=F) +
  geom_line(aes(y=mean)) +
  facet_wrap(~state) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2008-03-01','2008-11-03')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,1,0.02)) +
  labs(subtitle='pct_obama national')

natl_evs.2008.gg <-  ggplot(sim_evs, aes(x=t)) +
  geom_hline(yintercept = 270) +
  geom_line(aes(y=mean_dem_ev)) +
  geom_line(aes(y=median_dem_ev),linetype=2) +
  geom_ribbon(aes(ymin=low_dem_ev,ymax=high_dem_ev),alpha=0.2) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2008-03-01','2008-11-03')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,600,100)) +
  labs(subtitletitle='obama evs')

state_polls.2008.gg <- pct_obama %>%
  filter(state %in% ex_states) %>%
  left_join(df %>% select(state,t,pct_obama)) %>% # plot over time
  # plot
  ggplot(.,aes(x=t,col=state)) +
  geom_ribbon(aes(ymin=low,ymax=high),col=NA,alpha=0.2) +
  geom_hline(yintercept = 0.5) +
  geom_point(aes(y=pct_obama),alpha=0.2) +
  #geom_smooth(aes(y=pct_obama,col=state),method='loess',linetype=2,se=F) +
  geom_line(aes(y=mean)) +
  facet_wrap(~state) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2008-03-01','2008-11-03')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,1,0.05)) +
  labs(subtitle='pct_obama state')


# diff from national over time?
state_lean.2008.gg <- pct_obama[pct_obama$state != '--',] %>%
  left_join(pct_obama[pct_obama$state=='--',] %>%
              select(t,pct_obama_national=mean), by='t') %>%
  mutate(diff=mean-pct_obama_national) %>%
  group_by(state) %>%
  mutate(last_prob = last(prob)) %>%
  filter(state %in% ex_states) %>%
  ggplot(.,aes(x=t,y=diff,col=state)) +
  geom_hline(yintercept=0.0) +
  geom_line() +
  geom_label_repel(data = . %>% 
                     filter(t==max(t),
                            state %in% ex_states),
                   aes(label=state)) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2008-03-01','2008-11-03')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(-1,1,0.01)) +
  labs(subtitle = identifier)

# brier scores
# https://www.buzzfeednews.com/article/jsvine/2016-election-forecast-grades
compare <- pct_obama %>% 
  filter(t==max(t),state!='--') %>% 
  select(state,obama_win = prob) %>% 
  mutate(obama_win_actual = ifelse(state %in% c('CA','NV','OR','WA','CO','NM','MN','IL','VA','DC','MD','DE','NJ','CT','RI','MA','NH','VT','NY','HI','ME','MI','IA','OH','PA','WI','FL','NC','IN'),1,0),
         diff = (obama_win_actual - obama_win )^2) %>% 
  left_join(enframe(ev_state) %>% set_names(.,c('state','ev'))) %>% 
  mutate(ev_weight = ev/(sum(ev))) 


briers.2008 <- tibble(outlet='economist (backtest)',
       ev_wtd_brier = weighted.mean(compare$diff, compare$ev_weight),
       unwtd_brier = mean(compare$diff),
       states_correct=sum(round(compare$obama_win) == round(compare$obama_win_actual)))


# compare model estiamte v actual results
model_v_actual.2008 <- pct_obama %>% 
  filter(t==max(t),state!='--') %>% 
  left_join(politicaldata::pres_results %>% filter(year == 2008) %>%
              mutate(actual = dem/(dem+rep)) %>%
              dplyr::select(state,actual)) %>%
  filter(state != 'DC')

model_v_actual.2008.rmse <- sqrt(mean(c((model_v_actual.2008$mean - model_v_actual.2008$actual)^2)))

model_v_actual.2008.gg <- ggplot(model_v_actual.2008,aes(x=mean,y=actual,label=state)) +
  geom_abline() + 
  geom_text() + 
  stat_smooth(method='lm') +
  theme_minimal()

# model vs final polls vs prior
predictcompare.2008.gg <- pct_obama %>%
  filter(t == max(t),state %in% ex_states) %>%
  mutate(se = (high - mean)/1.68) %>%
  select(state,model_mean=mean,model_se=se) %>%
  left_join(df %>%
              filter(t > (max(t)-14),
                     state %in% ex_states) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_obama,n_respondents))) %>%
  left_join(enframe(mu_b_prior,'state','prior') %>%
              mutate(prior = inv.logit(prior))) %>%
  left_join(model_v_actual.2008 %>% dplyr::select(state,actual)) %>%
  ggplot(.,aes(y=state)) +
  geom_point(aes(x=poll,col='poll')) +
  geom_point(aes(x=model_mean,col='model')) +
  geom_point(aes(x=prior,col='prior')) +
  geom_point(aes(x=actual,col='result')) +
  theme_minimal()


# model better than polls?
 pct_obama %>%
  filter(t == max(t),state %in% ex_states) %>%
  mutate(se = (high - mean)/1.68) %>%
  select(state,model_mean=mean,model_se=se) %>%
  left_join(df %>%
              filter(t > (max(t)-14),
                     state %in% ex_states) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_obama,n_respondents))) %>%
  left_join(enframe(mu_b_prior,'state','prior') %>%
              mutate(prior = inv.logit(prior))) %>%
  left_join(model_v_actual.2008 %>% dplyr::select(state,actual)) %>% 
   summarise(poll_error = mean(abs(poll-actual)),
             model_error = mean(abs(model_mean-actual)))
 


# expected EVs roughly match the stochastic simulations?
final_states <- enframe(mu_b_prior,'state','prior') %>%
  left_join(df %>%
              filter(t > (max(t)-14),) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_obama,n_respondents))) %>%
  mutate(avg_poll_less_prior = mean(poll - inv.logit(prior), na.rm=T)) %>%
  mutate(poll = ifelse(is.na(poll),
                       inv.logit(prior) + avg_poll_less_prior,
                       poll)) %>%
  select(state,poll)

# final EV distribution
final_evs <- draws %>%
  left_join(state_elec_results %>% select(state,ev),by='state') %>%
  filter(t==max(t)) %>%
  group_by(draw) %>%
  summarise(dem_ev = sum(ev* (pct_obama > 0.5)))

final_evs.2008.gg <-  ggplot(final_evs,aes(x=dem_ev,
                       fill=ifelse(dem_ev>=270,'Democratic','Republican'))) +
    geom_vline(xintercept = 270) +
    geom_histogram(binwidth=1) +
    theme_minimal() +
    theme(legend.position = 'top',
          panel.grid.minor = element_blank()) +
    scale_fill_manual(name='Electoral College winner',values=c('Democratic'='#3A4EB1','Republican'='#E40A04')) +
    labs(x='Democratic electoral college votes',title='2008, Potential Electoral College outcomes',
         subtitle=sprintf("p(dem win) = %s | full stan model",round(mean(final_evs$dem_ev>=270),3)))


```

#### Map

```{r echo=F,message=F,warning=F}
map.2008.gg
```

#### Final electoral college histogram

```{r echo=F,message=F,warning=F}
final_evs.2008.gg
```

#### National and state polls and the electoral college over time

```{r echo=F,message=F,warning=F,fig.width = 9, fig.asp = .62}
grid.arrange(natl_polls.2008.gg, natl_evs.2008.gg, state_polls.2008.gg, 
             layout_matrix = rbind(c(1,1,3,3,3),
                                   c(2,2,3,3,3)),
             top = identifier
)
```

#### State vs national deltas over time

```{r echo=F,message=F,warning=F}
state_lean.2008.gg
```

#### Model results vs polls vs the prior

```{r echo=F,message=F,warning=F}
predictcompare.2008.gg
```

#### Performance

````{r echo=F,message=F,warning=F}
knitr::kable(briers.2008)
model_v_actual.2008.gg
model_v_actual.2008.rmse

```

#### Predictions for each state

```{r echo=F,message=F,warning=F}
pct_obama %>% 
  filter(t == max(t)) %>% 
  mutate(se = (high - mean)/1.68) %>%
  mutate(high = round(high,3),
         low = round(low,3),
         mean = round(mean,3),
         prob = round(prob,3),
         se = round(se,3)) %>%
  select(state,mean,low,high,prob,se) %>%
  arrange(abs(mean-0.5)) %>%
  datatable()
            
```

### 2012 

```{r include=F}
## Master variables
RUN_DATE <- min(ymd('2012-11-06'),Sys.Date())

election_day <- ymd("2012-11-06")
start_date <- as.Date("2012-03-01") # Keeping all polls after March 1, 2016
  
evs_df <- read_csv('data/2012.csv')

# will read in the most recent back-test for 2012
out  <- read_rds(sprintf('models/backtest_2012/stan_model_%s.rds',RUN_DATE))

# get all the polling data
#setwd(here("data/"))
all_polls <- read_csv('data/all_polls_2012.csv')

# select relevant columns from HufFPost polls
all_polls <- all_polls %>%
  dplyr::select(state, pollster, number.of.observations, mode,population,
                start.date, 
                end.date,
                obama, romney, undecided, other)%>%
  mutate(end.date = mdy(end.date),
         start.date = mdy(start.date))


# make sure we've got nothing from the futuree
all_polls <- all_polls %>%
  filter(ymd(end.date) <= RUN_DATE)

# basic mutations
df <- all_polls %>% 
  tbl_df %>%
  rename(n = number.of.observations) %>%
  mutate(begin = ymd(start.date),
         end   = ymd(end.date),
         t = end - (1 + as.numeric(end-begin)) %/% 2) %>%
  filter(t >= start_date & !is.na(t)
         & n > 1) 

# pollster mutations
df <- df %>%
  mutate(pollster = str_extract(pollster, pattern = "[A-z0-9 ]+") %>% sub("\\s+$", "", .),
         pollster = replace(pollster, pollster == "Fox News", "FOX"), # Fixing inconsistencies in pollster names
         pollster = replace(pollster, pollster == "WashPost", "Washington Post"),
         pollster = replace(pollster, pollster == "ABC News", "ABC"),
         undecided = ifelse(is.na(undecided), 0, undecided),
         other = ifelse(is.na(other), 0, other))

# vote shares etc
df <- df %>%
  mutate(two_party_sum = obama + romney,
         polltype = as.integer(as.character(recode(population, 
                                                   "Likely Voters" = "0", 
                                                   "Registered Voters" = "1",
                                                   "Adults" = "2"))), 
         n_respondents = round(n),
         # obama
         n_obama = round(n * obama/100),
         pct_obama = obama/two_party_sum,
         # romney
         n_romney = round(n * romney/100),
         pct_romney = romney/two_party_sum,
         # third-party
         n_other = round(n * other/100),
         p_other = other/100)


# state info

#here("data")
state_data <- read.csv("data/potus_results_76_16.csv")

all_state_names <- unique(state_data$state)

# Numerical indices passed to Stan for states, days, weeks, pollsters
df <- df %>% 
  mutate(poll_day = t - min(t) + 1,
         # Factors are alphabetically sorted: 1 = --, 2 = AL, 3 = AK, 4 = AZ...
         index_s = as.numeric(factor(as.character(state),
                                     levels = c('--',all_state_names))),  # ensure levels are same as all 50 names in sate_correlation
         index_t = 1 + as.numeric(t) - min(as.numeric(t)),
         index_p = as.numeric(as.factor(as.character(pollster))))  

T <- as.integer(round(difftime(election_day, min(df$start.date))))

# selections
df <- df %>%
  arrange(state, t, polltype, two_party_sum) %>% 
  distinct(state, t, pollster, .keep_all = TRUE) %>%
  select(
    # poll information
    state, t, begin, end, pollster, polltype, method = mode, n_respondents, 
    # vote shares
    pct_obama, n_obama, 
    pct_romney, n_romney, 
    p_other, n_other, poll_day, index_s, index_p, index_t) %>%
  mutate(index_s = ifelse(index_s == 1, 52, index_s - 1)) # national index = 51


# state weights
#setwd(here("data/"))
state_elec_results <- read.csv("data/2012.csv", 
                       header = TRUE, stringsAsFactors = FALSE) %>% 
  mutate(score = obama_count / (obama_count + romney_count),
         national_score = sum(obama_count)/sum(obama_count + romney_count),
         delta = score - national_score,
         share_national_vote = (total_count)
         /sum(total_count)) %>%
  arrange(state) 

rownames(state_elec_results) <- state_elec_results$state

# get state incdices
all_states <- state_elec_results$state
state_name <- state_elec_results$state_name
names(state_name) <- state_elec_results$state

# set prior differences
prior_diff_score <- state_elec_results$delta
names(prior_diff_score) <- all_states

# set state weights
state_weights <- c(state_elec_results$share_national_vote / sum(state_elec_results$share_national_vote))
names(state_weights) <- c(state_elec_results$state)

# electoral votes, by state:
ev_state <- state_elec_results$ev
names(ev_state) <- state_elec_results$state

# makes priors
national_mu_prior <- 0.502

# Mean of the mu_b_prior
mu_b_prior <- logit(national_mu_prior + prior_diff_score)

# etc
a <- rstan::extract(out, pars = "alpha")[[1]]
hist(a)
# sigmas
tibble(sigma_national = rstan::extract(out, pars = "sigma_a")[[1]],
       sigma_state = rstan::extract(out, pars = "sigma_b")[[1]]) %>%
  gather(parameter,value) %>%
  ggplot(.,aes(x=value)) +
  geom_histogram(binwidth=0.001) +
  facet_grid(rows=vars(parameter))
# measurement noise
measure_noise_national <- rstan::extract(out, pars = "measure_noise_national")[[1]] * mean(rstan::extract(out, pars = "sigma_measure_noise_national")[[1]])
hist(measure_noise_national)
# look at variation in mu_a
mu_a <- rstan::extract(out, pars = "mu_a")[[1]]
lapply(1:100,
       function(x){
         tibble(mu_a_draw = inv.logit(mu_a[x,]),
                trial = x) %>%
           mutate(date = min(df$end) + row_number()) 
       }) %>%
  do.call('bind_rows',.) %>%
  ggplot(.,aes(x=date,y=mu_a_draw,group=trial)) +
  geom_line(alpha=0.2)
# extract predictions
predicted_score <- rstan::extract(out, pars = "predicted_score")[[1]]

# states
pct_obama <- pblapply(1:dim(predicted_score)[3],
                    function(x){
                      # pred is mu_a + mu_b for the past, just mu_b for the future
                      temp <- predicted_score[,,x]
                      
                      # put in tibble
                      tibble(low = apply(temp,2,function(x){(quantile(x,0.05))}),
                             high = apply(temp,2,function(x){(quantile(x,0.95))}),
                             mean = apply(temp,2,function(x){(mean(x))}),
                             prob = apply(temp,2,function(x){(mean(x>0.5))}),
                             state = x) 
                      
                    }) %>% do.call('bind_rows',.)

pct_obama$state = colnames(state_correlation)[pct_obama$state]

pct_obama <- pct_obama %>%
  group_by(state) %>%
  mutate(t = row_number() + min(df$begin)) %>%
  ungroup()

# national
pct_obama_natl <- pblapply(1:dim(predicted_score)[2],
                         function(x){
                           # each row is a day for a particular draw
                           temp <- predicted_score[x,,] %>% as.data.frame()
                           names(temp) <- colnames(state_correlation)
                           
                           # for each row, get weigted natl vote
                           tibble(natl_vote = apply(temp,MARGIN = 1,function(y){weighted.mean(y,state_weights)})) %>%
                             mutate(t = row_number() + min(df$begin)) %>%
                             mutate(draw = x)
                         }) %>% do.call('bind_rows',.)

pct_obama_natl <- pct_obama_natl %>%
  group_by(t) %>%
  summarise(low = quantile(natl_vote,0.05),
            high = quantile(natl_vote,0.95),
            mean = mean(natl_vote),
            prob = mean(natl_vote > 0.5)) %>%
  mutate(state = '--')

# bind state and national vote
pct_obama <- pct_obama %>%
  bind_rows(pct_obama_natl) %>%
  arrange(desc(mean))

# look
ex_states <- c('IA','FL','OH','WI','MI','PA','AZ','NC','NH','TX','GA','MN')
pct_obama %>% filter(t == RUN_DATE,state %in% c(ex_states,'--')) %>% mutate(se = (high - mean)/1.68) %>% dplyr::select(-t)

map.2012.gg <- urbnmapr::states %>%
  left_join(pct_obama %>% filter(t == max(t)) %>%
              select(state_abbv=state,prob)) %>%
  ggplot(aes(x=long,y=lat,group=group,fill=prob)) +
  geom_polygon(col='gray40')  + 
  coord_map("albers",lat0=39, lat1=45) +
  scale_fill_gradient2(name='Democratic win probability',high='blue',low='red',mid='gray90',midpoint=0.5) +
  theme_void() + 
  theme(legend.position = 'top')


# electoral college by simulation
draws <- pblapply(1:dim(predicted_score)[3],
                  function(x){
                    # pred is mu_a + mu_b for the past, just mu_b for the future
                    pct_obama <- predicted_score[,,x]
                    
                    pct_obama <- pct_obama %>%
                      as.data.frame() %>%
                      mutate(draw = row_number()) %>%
                      gather(t,pct_obama,1:(ncol(.)-1)) %>%
                      mutate(t = as.numeric(gsub('V','',t)) + min(df$begin),
                             state = all_state_names[x]) 
                    
                    
                  }) %>% do.call('bind_rows',.)


sim_evs <- draws %>%
  left_join(state_elec_results %>% select(state,ev),by='state') %>%
  group_by(t,draw) %>%
  summarise(dem_ev = sum(ev * (pct_obama > 0.5))) %>%
  group_by(t) %>%
  summarise(mean_dem_ev = mean(dem_ev),
            median_dem_ev = median(dem_ev),
            high_dem_ev = quantile(dem_ev,0.975),
            low_dem_ev = quantile(dem_ev,0.025),
            prob = mean(dem_ev >= 270)) %>%
  left_join(pct_obama[pct_obama$state != '--',] %>%
              left_join(state_elec_results %>% select(state,ev),by='state') %>%
              group_by(t) %>%
              summarise(sum_dem_ev = sum(ev * (prob > 0.5))) )


# add identifier
identifier <- paste0(Sys.Date()," || " , out@model_name)

natl_polls.2012.gg <- pct_obama %>%
  filter(state == '--') %>%
  left_join(df %>% select(state,t,pct_obama)) %>% # plot over time
  # plot
  ggplot(.,aes(x=t)) +
  geom_ribbon(aes(ymin=low,ymax=high),col=NA,alpha=0.2) +
  geom_hline(yintercept = 0.5) +
  geom_hline(yintercept = national_mu_prior,linetype=2) +
  geom_point(aes(y=pct_obama),alpha=0.2) +
  geom_smooth(aes(y=pct_obama),method='loess',span=0.2,col='black',linetype=2,se=F) +
  geom_line(aes(y=mean)) +
  facet_wrap(~state) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2012-03-01','2012-11-06')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,1,0.02)) +
  labs(subtitle='pct_obama national')

natl_evs.2012.gg <-  ggplot(sim_evs, aes(x=t)) +
  geom_hline(yintercept = 270) +
  geom_line(aes(y=mean_dem_ev)) +
  geom_line(aes(y=median_dem_ev),linetype=2) +
  geom_ribbon(aes(ymin=low_dem_ev,ymax=high_dem_ev),alpha=0.2) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2012-03-01','2012-11-06')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,600,100)) +
  labs(subtitletitle='obama evs')

state_polls.2012.gg <- pct_obama %>%
  filter(state %in% ex_states) %>%
  left_join(df %>% select(state,t,pct_obama)) %>% # plot over time
  # plot
  ggplot(.,aes(x=t,col=state)) +
  geom_ribbon(aes(ymin=low,ymax=high),col=NA,alpha=0.2) +
  geom_hline(yintercept = 0.5) +
  geom_point(aes(y=pct_obama),alpha=0.2) +
  #geom_smooth(aes(y=pct_obama,col=state),method='loess',linetype=2,se=F) +
  geom_line(aes(y=mean)) +
  facet_wrap(~state) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2012-03-01','2012-11-06')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,1,0.05)) +
  labs(subtitle='pct_obama state')


# diff from national over time?
state_lean.2012.gg <- pct_obama[pct_obama$state != '--',] %>%
  left_join(pct_obama[pct_obama$state=='--',] %>%
              select(t,pct_obama_national=mean), by='t') %>%
  mutate(diff=mean-pct_obama_national) %>%
  group_by(state) %>%
  mutate(last_prob = last(prob)) %>%
  filter(state %in% ex_states) %>%
  ggplot(.,aes(x=t,y=diff,col=state)) +
  geom_hline(yintercept=0.0) +
  geom_line() +
  geom_label_repel(data = . %>% 
                     filter(t==max(t),
                            state %in% ex_states),
                   aes(label=state)) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2012-03-01','2012-11-06')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(-1,1,0.01)) +
  labs(subtitle = identifier)

# brier scores
# https://www.buzzfeednews.com/article/jsvine/2016-election-forecast-grades
compare <- pct_obama %>% 
  filter(t==max(t),state!='--') %>% 
  select(state,obama_win = prob) %>% 
  mutate(obama_win_actual = ifelse(state %in% c('CA','NV','OR','WA','CO','NM','MN','IL','VA','DC','MD','DE','NJ','CT','RI','MA','NH','VT','NY','HI','ME','MI','IA','OH','PA','WI','FL'),1,0),
         diff = (obama_win_actual - obama_win )^2) %>% 
  left_join(enframe(ev_state) %>% set_names(.,c('state','ev'))) %>% 
  mutate(ev_weight = ev/(sum(ev))) 


briers.2012 <- tibble(outlet = c('Linzer','Wang/Ferguson','Silver/538','Jackman/Pollster','Desart/Holbrook',
                  'Intrade','Enten/Margin of Error'),
       ev_wtd_brier = rep(NA,7),
       unwtd_brier = c(0.0038,0.00761,0.00911,0.00971,0.01605,0.02812,0.05075),
) %>% 
  bind_rows(tibble(outlet='economist (backtest)',
                   ev_wtd_brier = weighted.mean(compare$diff, compare$ev_weight),
                   unwtd_brier = mean(compare$diff),
                   states_correct=sum(round(compare$obama_win) == round(compare$obama_win_actual)))) %>%
  arrange(unwtd_brier)

# final v result
model_v_actual.2012 <- pct_obama %>% 
    filter(t==max(t),state!='--') %>% 
    left_join(politicaldata::pres_results %>% filter(year == 2012) %>%
                  mutate(actual = dem/(dem+rep)) %>%
                  dplyr::select(state,actual)) %>%
    filter(state != 'DC')

model_v_actual.2012.rmse <- sqrt(mean(c((model_v_actual.2012$mean - model_v_actual.2012$actual)^2)))

model_v_actual.2012.gg <- ggplot(model_v_actual.2012,aes(x=mean,y=actual,label=state)) +
    geom_abline() + 
    geom_text() + 
    stat_smooth(method='lm') +
    theme_minimal()


# model vs final polls vs prior
predictcompare.2012.gg <- pct_obama %>%
  filter(t == max(t),state %in% ex_states) %>%
  mutate(se = (high - mean)/1.68) %>%
  select(state,model_mean=mean,model_se=se) %>%
  left_join(df %>%
              filter(t > (max(t)-14),
                     state %in% ex_states) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_obama,n_respondents))) %>%
  left_join(enframe(mu_b_prior,'state','prior') %>%
              mutate(prior = inv.logit(prior))) %>%
  left_join(model_v_actual.2012 %>% dplyr::select(state,actual)) %>%
  ggplot(.,aes(y=state)) +
  geom_point(aes(x=poll,col='poll')) +
  geom_point(aes(x=model_mean,col='model')) +
  geom_point(aes(x=prior,col='prior')) +
  geom_point(aes(x=actual,col='result')) +
  theme_minimal()


# model better than polls?
 pct_obama %>%
  filter(t == max(t),state %in% ex_states) %>%
  mutate(se = (high - mean)/1.68) %>%
  select(state,model_mean=mean,model_se=se) %>%
  left_join(df %>%
              filter(t > (max(t)-14),
                     state %in% ex_states) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_obama,n_respondents))) %>%
  left_join(enframe(mu_b_prior,'state','prior') %>%
              mutate(prior = inv.logit(prior))) %>%
  left_join(model_v_actual.2012 %>% dplyr::select(state,actual)) %>% 
   summarise(poll_error = mean(abs(poll-actual)),
             model_error = mean(abs(model_mean-actual)))
 
# expected EVs roughly match the stochastic simulations?
final_states <- enframe(mu_b_prior,'state','prior') %>%
  left_join(df %>%
              filter(t > (max(t)-14),) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_obama,n_respondents))) %>%
  mutate(avg_poll_less_prior = mean(poll - inv.logit(prior), na.rm=T)) %>%
  mutate(poll = ifelse(is.na(poll),
                       inv.logit(prior) + avg_poll_less_prior,
                       poll)) %>%
  select(state,poll)

# final EV distribution
final_evs <- draws %>%
  left_join(state_elec_results %>% select(state,ev),by='state') %>%
  filter(t==max(t)) %>%
  group_by(draw) %>%
  summarise(dem_ev = sum(ev* (pct_obama > 0.5)))

final_evs.2012.gg <-  ggplot(final_evs,aes(x=dem_ev,
                       fill=ifelse(dem_ev>=270,'Democratic','Republican'))) +
    geom_vline(xintercept = 270) +
    geom_histogram(binwidth=1) +
    theme_minimal() +
    theme(legend.position = 'top',
          panel.grid.minor = element_blank()) +
    scale_fill_manual(name='Electoral College winner',values=c('Democratic'='#3A4EB1','Republican'='#E40A04')) +
    labs(x='Democratic electoral college votes',title='2012, Potential Electoral College outcomes',
         subtitle=sprintf("p(dem win) = %s | full stan model",round(mean(final_evs$dem_ev>=270),3)))


```


#### Map

```{r echo=F,message=F,warning=F}
map.2012.gg
```

#### Final electoral college histogram

```{r echo=F,message=F,warning=F}
final_evs.2012.gg
```

#### National and state polls and the electoral college over time

```{r echo=F,message=F,warning=F,fig.width = 9, fig.asp = .62}
grid.arrange(natl_polls.2012.gg, natl_evs.2012.gg, state_polls.2012.gg, 
             layout_matrix = rbind(c(1,1,3,3,3),
                                   c(2,2,3,3,3)),
             top = identifier
)
```

#### State vs national deltas over time

```{r echo=F,message=F,warning=F}
state_lean.2012.gg
```

#### Model results vs polls vs the prior

```{r echo=F,message=F,warning=F}
predictcompare.2012.gg
```

#### Performance

````{r echo=F,message=F,warning=F}

knitr::kable(briers.2012)
model_v_actual.2012.gg
model_v_actual.2012.rmse
```


#### Predictions for each state

```{r echo=F,message=F,warning=F}
pct_obama %>% 
  filter(t == max(t)) %>% 
  mutate(se = (high - mean)/1.68) %>%
  mutate(high = round(high,3),
         low = round(low,3),
         mean = round(mean,3),
         prob = round(prob,3),
         se = round(se,3)) %>%
  select(state,mean,low,high,prob,se) %>%
  arrange(abs(mean-0.5)) %>%
  datatable()
            
```

### 2016 

```{r include=F}
## Master variables
RUN_DATE <- min(ymd('2016-11-08'),Sys.Date())

election_day <- ymd("2016-11-08")
start_date <- as.Date("2016-03-01") # Keeping all polls after March 1, 2016
  
evs_df <- read_csv('data/2012.csv')

# will read in the most recent back-test for 2012
out  <- read_rds(sprintf('models/stan_model_%s.rds',RUN_DATE))

# get all the polling data
#setwd(here("data/"))
all_polls <- read_csv('data/all_polls.csv')

# select relevant columns from HufFPost polls
all_polls <- all_polls %>%
  dplyr::select(state, pollster, number.of.observations, mode,population,
                start.date, 
                end.date,
                clinton, trump, undecided, other)%>%
  mutate(end.date = ymd(end.date),
         start.date = ymd(start.date))


# make sure we've got nothing from the futuree
all_polls <- all_polls %>%
  filter(ymd(end.date) <= RUN_DATE)

# basic mutations
df <- all_polls %>% 
  tbl_df %>%
  rename(n = number.of.observations) %>%
  mutate(begin = ymd(start.date),
         end   = ymd(end.date),
         t = end - (1 + as.numeric(end-begin)) %/% 2) %>%
  filter(t >= start_date & !is.na(t)
         & n > 1) 

# pollster mutations
df <- df %>%
  mutate(pollster = str_extract(pollster, pattern = "[A-z0-9 ]+") %>% sub("\\s+$", "", .),
         pollster = replace(pollster, pollster == "Fox News", "FOX"), # Fixing inconsistencies in pollster names
         pollster = replace(pollster, pollster == "WashPost", "Washington Post"),
         pollster = replace(pollster, pollster == "ABC News", "ABC"),
         undecided = ifelse(is.na(undecided), 0, undecided),
         other = ifelse(is.na(other), 0, other))

# vote shares etc
df <- df %>%
  mutate(two_party_sum = clinton + trump,
         polltype = as.integer(as.character(recode(population, 
                                                   "Likely Voters" = "0", 
                                                   "Registered Voters" = "1",
                                                   "Adults" = "2"))), 
         n_respondents = round(n),
         # clinton
         n_clinton = round(n * clinton/100),
         pct_clinton = clinton/two_party_sum,
         # trump
         n_trump = round(n * trump/100),
         pct_trump = trump/two_party_sum,
         # third-party
         n_other = round(n * other/100),
         p_other = other/100)


# state info
state_correlation <- read.csv("data/potus_results_76_16.csv") %>% 
    select(year, state, dem) %>%
    group_by(state) %>%
    mutate(dem = dem )  %>% 
    spread(state,dem) %>% 
    dplyr::select(-year)

#here("data")
state_data <- read.csv("data/potus_results_76_16.csv")

all_state_names <- unique(state_data$state)

# Numerical indices passed to Stan for states, days, weeks, pollsters
df <- df %>% 
  mutate(poll_day = t - min(t) + 1,
         # Factors are alphabetically sorted: 1 = --, 2 = AL, 3 = AK, 4 = AZ...
         index_s = as.numeric(factor(as.character(state),
                                     levels = c('--',all_state_names))),  # ensure levels are same as all 50 names in sate_correlation
         index_t = 1 + as.numeric(t) - min(as.numeric(t)),
         index_p = as.numeric(as.factor(as.character(pollster))))  

T <- as.integer(round(difftime(election_day, min(df$start.date))))

# selections
df <- df %>%
  arrange(state, t, polltype, two_party_sum) %>% 
  distinct(state, t, pollster, .keep_all = TRUE) %>%
  select(
    # poll information
    state, t, begin, end, pollster, polltype, method = mode, n_respondents, 
    # vote shares
    pct_clinton, n_clinton, 
    pct_trump, n_trump, 
    p_other, n_other, poll_day, index_s, index_p, index_t) %>%
  mutate(index_s = ifelse(index_s == 1, 52, index_s - 1)) # national index = 51


# state weights
#setwd(here("data/"))
state_elec_results <- read.csv("data/2012.csv", 
                       header = TRUE, stringsAsFactors = FALSE) %>% 
  mutate(score = obama_count / (obama_count + romney_count),
         national_score = sum(obama_count)/sum(obama_count + romney_count),
         delta = score - national_score,
         share_national_vote = (total_count)
         /sum(total_count)) %>%
  arrange(state) 

rownames(state_elec_results) <- state_elec_results$state

# get state incdices
all_states <- state_elec_results$state
state_name <- state_elec_results$state_name
names(state_name) <- state_elec_results$state

# set prior differences
prior_diff_score <- state_elec_results$delta
names(prior_diff_score) <- all_states

# set state weights
state_weights <- c(state_elec_results$share_national_vote / sum(state_elec_results$share_national_vote))
names(state_weights) <- c(state_elec_results$state)

# electoral votes, by state:
ev_state <- state_elec_results$ev
names(ev_state) <- state_elec_results$state

# makes priors
national_mu_prior <- 0.506

# Mean of the mu_b_prior
mu_b_prior <- logit(national_mu_prior + prior_diff_score)

# etc
a <- rstan::extract(out, pars = "alpha")[[1]]
hist(a)
# sigmas
tibble(sigma_national = rstan::extract(out, pars = "sigma_a")[[1]],
       sigma_state = rstan::extract(out, pars = "sigma_b")[[1]]) %>%
  gather(parameter,value) %>%
  ggplot(.,aes(x=value)) +
  geom_histogram(binwidth=0.001) +
  facet_grid(rows=vars(parameter))
# measurement noise
measure_noise_national <- rstan::extract(out, pars = "measure_noise_national")[[1]] * mean(rstan::extract(out, pars = "sigma_measure_noise_national")[[1]])
hist(measure_noise_national)
# look at variation in mu_a
mu_a <- rstan::extract(out, pars = "mu_a")[[1]]
lapply(1:100,
       function(x){
         tibble(mu_a_draw = inv.logit(mu_a[x,]),
                trial = x) %>%
           mutate(date = min(df$end) + row_number()) 
       }) %>%
  do.call('bind_rows',.) %>%
  ggplot(.,aes(x=date,y=mu_a_draw,group=trial)) +
  geom_line(alpha=0.2)
# extract predictions
predicted_score <- rstan::extract(out, pars = "predicted_score")[[1]]

# states
pct_clinton <- pblapply(1:dim(predicted_score)[3],
                    function(x){
                      # pred is mu_a + mu_b for the past, just mu_b for the future
                      temp <- predicted_score[,,x]
                      
                      # put in tibble
                      tibble(low = apply(temp,2,function(x){(quantile(x,0.05))}),
                             high = apply(temp,2,function(x){(quantile(x,0.95))}),
                             mean = apply(temp,2,function(x){(mean(x))}),
                             prob = apply(temp,2,function(x){(mean(x>0.5))}),
                             state = x) 
                      
                    }) %>% do.call('bind_rows',.)

pct_clinton$state = colnames(state_correlation)[pct_clinton$state]

pct_clinton <- pct_clinton %>%
  group_by(state) %>%
  mutate(t = row_number() + min(df$begin)) %>%
  ungroup()

# national
pct_clinton_natl <- pblapply(1:dim(predicted_score)[2],
                         function(x){
                           # each row is a day for a particular draw
                           temp <- predicted_score[x,,] %>% as.data.frame()
                           names(temp) <- colnames(state_correlation)
                           
                           # for each row, get weigted natl vote
                           tibble(natl_vote = apply(temp,MARGIN = 1,function(y){weighted.mean(y,state_weights)})) %>%
                             mutate(t = row_number() + min(df$begin)) %>%
                             mutate(draw = x)
                         }) %>% do.call('bind_rows',.)

pct_clinton_natl <- pct_clinton_natl %>%
  group_by(t) %>%
  summarise(low = quantile(natl_vote,0.05),
            high = quantile(natl_vote,0.95),
            mean = mean(natl_vote),
            prob = mean(natl_vote > 0.5)) %>%
  mutate(state = '--')

# bind state and national vote
pct_clinton <- pct_clinton %>%
  bind_rows(pct_clinton_natl) %>%
  arrange(desc(mean))

# look
ex_states <- c('IA','FL','OH','WI','MI','PA','AZ','NC','NH','TX','GA','MN')
pct_clinton %>% filter(t == RUN_DATE,state %in% c(ex_states,'--')) %>% mutate(se = (high - mean)/1.68) %>% dplyr::select(-t)

map.2016.gg <- urbnmapr::states %>%
  left_join(pct_clinton %>% filter(t == max(t)) %>%
              select(state_abbv=state,prob)) %>%
  ggplot(aes(x=long,y=lat,group=group,fill=prob)) +
  geom_polygon(col='gray40')  + 
  coord_map("albers",lat0=39, lat1=45) +
  scale_fill_gradient2(name='Democratic win probability',high='blue',low='red',mid='gray90',midpoint=0.5) +
  theme_void() + 
  theme(legend.position = 'top')


# electoral college by simulation
draws <- pblapply(1:dim(predicted_score)[3],
                  function(x){
                    # pred is mu_a + mu_b for the past, just mu_b for the future
                    pct_clinton <- predicted_score[,,x]
                    
                    pct_clinton <- pct_clinton %>%
                      as.data.frame() %>%
                      mutate(draw = row_number()) %>%
                      gather(t,pct_clinton,1:(ncol(.)-1)) %>%
                      mutate(t = as.numeric(gsub('V','',t)) + min(df$begin),
                             state = all_state_names[x]) 
                    
                    
                  }) %>% do.call('bind_rows',.)


sim_evs <- draws %>%
  left_join(state_elec_results %>% select(state,ev),by='state') %>%
  group_by(t,draw) %>%
  summarise(dem_ev = sum(ev * (pct_clinton > 0.5))) %>%
  group_by(t) %>%
  summarise(mean_dem_ev = mean(dem_ev),
            median_dem_ev = median(dem_ev),
            high_dem_ev = quantile(dem_ev,0.975),
            low_dem_ev = quantile(dem_ev,0.025),
            prob = mean(dem_ev >= 270)) %>%
  left_join(pct_clinton[pct_clinton$state != '--',] %>%
              left_join(state_elec_results %>% select(state,ev),by='state') %>%
              group_by(t) %>%
              summarise(sum_dem_ev = sum(ev * (prob > 0.5))) )


# add identifier
identifier <- paste0(Sys.Date()," || " , out@model_name)

natl_polls.2016.gg <- pct_clinton %>%
  filter(state == '--') %>%
  left_join(df %>% select(state,t,pct_clinton)) %>% # plot over time
  # plot
  ggplot(.,aes(x=t)) +
  geom_ribbon(aes(ymin=low,ymax=high),col=NA,alpha=0.2) +
  geom_hline(yintercept = 0.5) +
  geom_hline(yintercept = national_mu_prior,linetype=2) +
  geom_point(aes(y=pct_clinton),alpha=0.2) +
  geom_smooth(aes(y=pct_clinton),method='loess',span=0.2,col='black',linetype=2,se=F) +
  geom_line(aes(y=mean)) +
  facet_wrap(~state) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2016-03-01','2016-11-08')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,1,0.02)) +
  labs(subtitle='pct_clinton national')

natl_evs.2016.gg <-  ggplot(sim_evs, aes(x=t)) +
  geom_hline(yintercept = 270) +
  geom_line(aes(y=mean_dem_ev)) +
  geom_line(aes(y=median_dem_ev),linetype=2) +
  geom_ribbon(aes(ymin=low_dem_ev,ymax=high_dem_ev),alpha=0.2) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2016-03-01','2016-11-08')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,600,100)) +
  labs(subtitletitle='clinton evs')

state_polls.2016.gg <- pct_clinton %>%
  filter(state %in% ex_states) %>%
  left_join(df %>% select(state,t,pct_clinton)) %>% # plot over time
  # plot
  ggplot(.,aes(x=t,col=state)) +
  geom_ribbon(aes(ymin=low,ymax=high),col=NA,alpha=0.2) +
  geom_hline(yintercept = 0.5) +
  geom_point(aes(y=pct_clinton),alpha=0.2) +
  #geom_smooth(aes(y=pct_clinton,col=state),method='loess',linetype=2,se=F) +
  geom_line(aes(y=mean)) +
  facet_wrap(~state) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2016-03-01','2016-11-08')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(0,1,0.05)) +
  labs(subtitle='pct_clinton state')


# diff from national over time?
state_lean.2016.gg <- pct_clinton[pct_clinton$state != '--',] %>%
  left_join(pct_clinton[pct_clinton$state=='--',] %>%
              select(t,pct_clinton_national=mean), by='t') %>%
  mutate(diff=mean-pct_clinton_national) %>%
  group_by(state) %>%
  mutate(last_prob = last(prob)) %>%
  filter(state %in% ex_states) %>%
  ggplot(.,aes(x=t,y=diff,col=state)) +
  geom_hline(yintercept=0.0) +
  geom_line() +
  geom_label_repel(data = . %>% 
                     filter(t==max(t),
                            state %in% ex_states),
                   aes(label=state)) +
  theme_minimal()  +
  theme(legend.position = 'none') +
  scale_x_date(limits=c(ymd('2016-03-01','2016-11-08')),date_breaks='1 month',date_labels='%b') +
  scale_y_continuous(breaks=seq(-1,1,0.01)) +
  labs(subtitle = identifier)

# brier scores
# https://www.buzzfeednews.com/article/jsvine/2016-election-forecast-grades
compare <- pct_clinton %>% 
  filter(t==max(t),state!='--') %>% 
  select(state,clinton_win=prob) %>% 
  mutate(clinton_win_actual = ifelse(state %in% c('CA','NV','OR','WA','CO','NM','MN','IL','VA','DC','MD','DE','NJ','CT','RI','MA','NH','VT','NY','HI','ME'),1,0),
         diff = (clinton_win_actual - clinton_win )^2) %>% left_join(enframe(ev_state) %>% set_names(.,c('state','ev'))) %>% 
  mutate(ev_weight = ev/(sum(ev))) 


briers.2016 <- tibble(outlet = c('538 polls-plus','538 polls-only','princeton','nyt upshot','kremp/slate','pollsavvy','predictwise markets','predictwise overall','desart and holbrook','daily kos','huffpost'),
       ev_wtd_brier = c(0.0928,0.0936,0.1169,0.1208,0.121,0.1219,0.1272,0.1276,0.1279,0.1439,0.1505),
       unwtd_brier = c(0.0664,0.0672,0.0744,0.0801,0.0766,0.0794,0.0767,0.0783,0.0825,0.0864,0.0892),
       states_correct = c(46,46,47,46,46,46,46,46,44,46,46)) %>% 
  bind_rows(tibble(outlet='economist (backtest)',
                   ev_wtd_brier = weighted.mean(compare$diff, compare$ev_weight),
                   unwtd_brier = mean(compare$diff),
                   states_correct=sum(round(compare$clinton_win) == round(compare$clinton_win_actual)))) %>%
  arrange(ev_wtd_brier) 


# compare model estiamte v actual results
model_v_actual.2016 <- pct_clinton %>% 
  filter(t==max(t),state!='--') %>% 
  left_join(politicaldata::pres_results %>% filter(year == 2016) %>%
              mutate(actual = dem/(dem+rep)) %>%
              dplyr::select(state,actual)) %>%
  filter(state != 'DC') 

model_v_actual.2016.rmse <- sqrt(mean(c((model_v_actual.2016$mean - model_v_actual.2016$actual)^2)))

model_v_actual.2016.gg <- ggplot(model_v_actual.2016,aes(x=mean,y=actual,label=state)) +
  geom_abline() + 
  geom_text() + 
  stat_smooth(method='lm') +
  theme_minimal()

# model vs final polls vs prior
predictcompare.2016.gg <- pct_clinton %>%
  filter(t == max(t),state %in% ex_states) %>%
  mutate(se = (high - mean)/1.68) %>%
  select(state,model_mean=mean,model_se=se) %>%
  left_join(df %>%
              filter(t > (max(t)-14),
                     state %in% ex_states) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_clinton,n_respondents))) %>%
  left_join(enframe(mu_b_prior,'state','prior') %>%
              mutate(prior = inv.logit(prior))) %>%
  left_join(model_v_actual.2016 %>% dplyr::select(state,actual)) %>%
  ggplot(.,aes(y=state)) +
  geom_point(aes(x=poll,col='poll')) +
  geom_point(aes(x=model_mean,col='model')) +
  geom_point(aes(x=prior,col='prior')) +
  geom_point(aes(x=actual,col='result')) +
  theme_minimal()


# model better than polls?
 pct_clinton %>%
  filter(t == max(t),state %in% ex_states) %>%
  mutate(se = (high - mean)/1.68) %>%
  select(state,model_mean=mean,model_se=se) %>%
  left_join(df %>%
              filter(t > (max(t)-14),
                     state %in% ex_states) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_clinton,n_respondents))) %>%
  left_join(enframe(mu_b_prior,'state','prior') %>%
              mutate(prior = inv.logit(prior))) %>%
  left_join(model_v_actual.2016 %>% dplyr::select(state,actual)) %>% 
   summarise(poll_error = mean(abs(poll-actual)),
             model_error = mean(abs(model_mean-actual)))

# expected EVs roughly match the stochastic simulations?
final_states <- enframe(mu_b_prior,'state','prior') %>%
  left_join(df %>%
              filter(t > (max(t)-14),) %>%
              group_by(state) %>%
              summarise(poll = weighted.mean(pct_clinton,n_respondents))) %>%
  mutate(avg_poll_less_prior = mean(poll - inv.logit(prior), na.rm=T)) %>%
  mutate(poll = ifelse(is.na(poll),
                       inv.logit(prior) + avg_poll_less_prior,
                       poll)) %>%
  select(state,poll)

# final EV distribution
final_evs <- draws %>%
  left_join(state_elec_results %>% select(state,ev),by='state') %>%
  filter(t==max(t)) %>%
  group_by(draw) %>%
  summarise(dem_ev = sum(ev* (pct_clinton > 0.5)))

final_evs.2016.gg <-  ggplot(final_evs,aes(x=dem_ev,
                       fill=ifelse(dem_ev>=270,'Democratic','Republican'))) +
    geom_vline(xintercept = 270) +
    geom_histogram(binwidth=1) +
    theme_minimal() +
    theme(legend.position = 'top',
          panel.grid.minor = element_blank()) +
    scale_fill_manual(name='Electoral College winner',values=c('Democratic'='#3A4EB1','Republican'='#E40A04')) +
    labs(x='Democratic electoral college votes',title='2016, Potential Electoral College outcomes',
         subtitle=sprintf("p(dem win) = %s | full stan model",round(mean(final_evs$dem_ev>=270),3)))


```


#### Map

```{r echo=F,message=F,warning=F}
map.2016.gg
```

#### Final electoral college histogram

```{r echo=F,message=F,warning=F}
final_evs.2016.gg
```

#### National and state polls and the electoral college over time

```{r echo=F,message=F,warning=F,fig.width = 9, fig.asp = .62}
grid.arrange(natl_polls.2016.gg, natl_evs.2016.gg, state_polls.2016.gg, 
             layout_matrix = rbind(c(1,1,3,3,3),
                                   c(2,2,3,3,3)),
             top = identifier
)
```

#### State vs national deltas over time

```{r echo=F,message=F,warning=F}
state_lean.2016.gg
```

#### Model results vs polls vs the prior

```{r echo=F,message=F,warning=F}
predictcompare.2016.gg
```

#### Performance

````{r echo=F,message=F,warning=F}
knitr::kable(briers.2016)
model_v_actual.2016.gg
model_v_actual.2016.rmse
```

#### Predictions for each state

```{r echo=F,message=F,warning=F}
pct_clinton %>% 
  filter(t == max(t)) %>% 
  mutate(se = (high - mean)/1.68) %>%
  mutate(high = round(high,3),
         low = round(low,3),
         mean = round(mean,3),
         prob = round(prob,3),
         se = round(se,3)) %>%
  select(state,mean,low,high,prob,se) %>%
  arrange(abs(mean-0.5)) %>%
  datatable()
            
```



## Cumulative charts

### Calibration plot

```{r echo=F,message=F,warning=F}
calibration_data <- model_v_actual.2016 %>% mutate(year = 2016) %>%
  bind_rows(model_v_actual.2012 %>% mutate(year = 2012)) %>%
  bind_rows(model_v_actual.2008 %>% mutate(year = 2008))

calibration.gg <- ggplot(calibration_data, aes(x=prob,y=ifelse(actual>0.5,1,0),
                                               label = paste0(state,'-',year))) + 
  geom_point() + 
  geom_text_repel(data = calibration_data %>%
                    filter(prob < 0.6 & actual>0.5 | 
                                    prob > 0.4 & actual<0.5 )) + 
  geom_smooth() +
  geom_abline()  +
  scale_x_continuous(breaks=seq(0,1,0.1)) +
  scale_y_continuous(breaks=seq(0,1,0.1)) +
  theme_minimal()

print(calibration.gg)

```


